Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-17 main.tex
Added equipment specs, fixes #10.

\section{Methods}
%
\subsection{Bicycle}
%
We modified an electric bicycle (Grenoble C8 HMB, Royal Dutch Gazelle, Dieren,
The Netherlands) with a custom motor in the steering assembly capable of
applying up to 7~\si{\newton\meter} of torque between the head tube and steer
tube, see Figure~\ref{fig:balance-assist-bicycle}. A custom motor controller
converts the commanded torque to applied torque. We measure the speed of the
rear wheel with an encoder (Magura XXX) and measure the body fixed roll rate of
the bicycle with a MEMs rate gyroscope (BN0086, Sparkfun, Niwot, USA). The
balance assist control algorithm is implemented on a microprocessor (Teensy,
PJRC, USA) and data from all sensors is logged with a CAN bus (CanEdge2, CSS
Electronics, Denmark) at at least 100 Hz.

@@

\subsection{Perturbations}
%
We apply longitudinal forces to the ends of each handlebar from Earth fixed
posts using an adapted Bump'Em system~\citep{Tan2020} which we arrange with
four motors working in cooperation, Figure~\ref{fig:setup-diagram}. The four
motors are programmed to apply a light force at all times to keep the ropes
taught and to track a commanded force profile using a PID controller running on
a microprocessor (Arduino Mega 2560, Arduino, Italy). We measure the force
applied by each motor at the handlebar via four inline 250~\si{\newton} S-style
load cells except the left rear load cell that has a max load of
500~\si{\newton}. The commanded force profiles are designed to apply an
external pulsive torque to the front assembly (handlebars, forks, wheel) at
magnitudes varying from \SIrange{16}{160}{\newton\meter}. The four motors
(EC-90 Flat, Maxon Group, Switzerland) are arranged at the four corners of a
1~\si{\meter} wide treadmill that can reach speeds of
18~\si{\kilo\meter\per\hour}. The general base design of this system is
described in detail in Van De Velde's MSc thesis~\citep{vandeVelde2022a} and
the physical arrangement is shown in Figure~\ref{fig:participant-in-set-up}.
Our modifications relative to Van De Velde include simplifying the controller
with a inexpensive microcontroller and the use of a simpler non-active safety
harness.
%
% NOTE : I get a compile error without the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}

>>>>

\section{Methods}
%
\subsection{Bicycle}
%
We modified an electric bicycle (Grenoble C8 HMB, Royal Dutch Gazelle, Dieren,
The Netherlands) with a custom motor in the steering assembly capable of
applying up to 7~\si{\newton\meter} of torque between the head tube and steer
tube, see Figure~\ref{fig:balance-assist-bicycle}. A custom motor controller
converts the commanded torque to applied torque. We measure the speed of the
rear wheel with an encoder (ABS Speed Sensor, Magura GmbH, Bad Urach, Germany)
and measure the body fixed roll rate of the bicycle with a MEMs rate gyroscope
(VR IMU BN0086, Sparkfun, Niwot, USA). The balance assist control algorithm is
implemented on a microprocessor (Teensy, PJRC, USA) and data from all sensors
is logged with a CAN bus (CanEdge2, CSS Electronics, Aabyhoej, Denmark) at at
least 100 Hz.

@@

\subsection{Perturbations}
%
We apply longitudinal forces to the ends of each handlebar from Earth fixed
posts using an adapted Bump'Em system~\citep{Tan2020} which we arrange with
four motors working in cooperation, Figure~\ref{fig:setup-diagram}. The four
motors are programmed to apply a light force at all times to keep the ropes
taught and to track a commanded force profile using a PID controller running on
a microprocessor (Arduino Mega 2560, Arduino, Italy). We measure the force
applied by each motor at the handlebar via four inline 250~\si{\newton} S-style
load cells (Bosche GmbH, Damme, Germany) except the left rear load cell that
has a max load of 500~\si{\newton}. The commanded force profiles are designed
to apply an external pulsive torque to the front assembly (handlebars, forks,
wheel) at magnitudes varying from \SIrange{16}{160}{\newton\meter}. The four
motors (EC-90 Flat, Maxon Group, Switzerland) are arranged at the four corners
of a 1~\si{\meter} wide treadmill (Bonte Technology B.V., Zwolle, The
Netherlands) that can reach speeds of 18~\si{\kilo\meter\per\hour}. The general
design of this system is described in detail in Van De Velde's MSc
thesis~\citep{vandeVelde2022a} and the physical arrangement is shown in
Figure~\ref{fig:participant-in-set-up}. Our modifications relative to Van De
Velde include simplifying the Bump'Em motor controller with an inexpensive
microcontroller and the use of a simpler non-active safety harness.
%
% NOTE : I get a compile error without the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}
